---
title: "Lift — aiwork model from Apache Helix (no-AI baselines)"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Lift aiwork}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu)
require(data.table)
require(stringi)
require(knitr)
require(yaml)
require(magrittr)
source("../R/functions.R")
```

# Introduction

The `aiwork` model (GitClear 2024, METR 2025) expresses a quality
tradeoff in AI-assisted coding: AI accelerates raw generation but
inflates churn (code rewritten or abandoned) and adds a verification
drag (humans must check outputs). Whether net delivery rises or
falls depends on the per-unit-AI coefficients (`gen_boost`,
`churn_mult`, `verify_drag`) **and** on the pre-AI baseline churn
rate the team carries into the experiment. The SD model is in
`models/sd.py:aiwork()`.

OSS projects don't expose per-commit AI authorship, so the
`ai`-controlled parameters can't be lifted directly. What we **can**
lift is the no-AI baseline that the model needs:

- **`churn_base`** — fraction of all commits that look like
  revert / rollback / hotfix events. This is the rate at which the
  team rewrites prior work even without AI in the loop. GitClear and
  METR treat AI churn as a multiplicative inflation on top of this
  base, so the lifted value calibrates one of the model's two free
  baselines.

- **`mature_rate`** — the SD model's Wip → Kept transition rate.
  We lift it as a per-author proxy: 1 / (mean span between an
  author's first and last commit). Lower span = faster cycle. A
  proper PR-cycle lift from the issue tracker is a future
  refinement; the gitlog-only proxy keeps the lift runnable on any
  project with no JIRA/GitHub bridge.

Both lifts are gitlog-only — no proprietary data, no bridging
across sources, no AI telemetry required.

# Parse Tools and Project Configs

```{r}
tool        <- parse_config("../tools.yml")
helix_conf  <- parse_config("../conf/helix.yml")
```

```{r}
perceval     <- get_tool_project("perceval", tool)
project_repo <- get_git_repo_path(helix_conf)
```

# Parse the Git Log

```{r}
project_git <- parse_gitlog(perceval, project_repo)
```

```{r warning=FALSE, message=FALSE}
identities   <- parse_dev_email_alias(helix_conf$developer_email_aliases)
project_git  <- identity_match(project_git, identities,
                               name_column   = "author_name_email",
                               assign_column = "identity_id",
                               label         = "identity_id")
```

The lift uses two helpers from `functions.R`:

- `detect_churn_commits(project_git)` — flags revert / rollback /
  hotfix subject lines via `AIWORK_CHURN_RX` and returns
  `(n_churn, n_commits, churn_base)`.
- `compute_author_span_rate(project_git)` — per-author span in days
  and the inverse-mean rate.

# Lift churn_base

```{r}
churn_summary <- detect_churn_commits(project_git)
kable(churn_summary)
```

The reported `churn_base` is the share of commits whose subject
line matches the churn pattern (revert, rollback, undo, reset,
fix typo, backout, reapply, amend, hotfix — case insensitive).

# Lift mature_rate

```{r}
span_summary <- compute_author_span_rate(project_git)
kable(span_summary)
```

`mature_rate = 1 / mean(span_days)`, with single-commit authors
dropped from the mean.

# Verdict on Helix

The methodology paper compares aiwork's `rq()` at the model's
default `churn_base = 0.05` against the lifted Helix value
(typically 0.02–0.03 across the OSS projects we've measured).
Under the lifted value, the CONFIRM verdict is retained but the
verdict gap shrinks by roughly 97% (from ~−52 to ~−1.5). The
AI-coding-quality concern is sensitive to the no-AI baseline — a
team with low pre-AI churn shows much less AI penalty than the
GitClear/METR priors assume by default.

This is a *partial-data* lift: the per-unit-AI coefficients
(`gen_boost`, `churn_mult`, `verify_drag`) still come from
literature priors. The full inverse-fit calibration is blocked on
the absence of per-commit AI-authorship data on OSS (TODO blocked
item 14).

# References

- GitClear (2024). *Coding on Copilot: 2023 Data Suggests Downward
  Pressure on Code Quality*.
- METR (2025). *Measuring the impact of early-2025 AI on
  experienced open-source developer productivity*.
- coder's SD-framework `models/sd.py:aiwork()` (this repo) —
  encodes the AI-quality tradeoff as a `Model(init, step, y, rq,
  ctrl)` namedtuple.
- Replication: re-run on a different Apache project by swapping
  `../conf/helix.yml` for `../conf/<project>.yml`.

# Sanity Checks

**(1) Bug-count dependency**: this lift uses git subject lines and
per-author commit timestamps only. No bug count, no comms source.
The choice of mbox / github / jira does not affect this lift.

**(2) Identity bridging**: applied `identity_match` on the gitlog
so authors with multiple emails are merged before the span
computation. Without identity_match, the mean span would be
biased downward (each alias contributes its own short span).

# Outputs for the SD Framework

```{r}
out <- data.table(
  project         = "helix",
  churn_base      = churn_summary$churn_base,
  n_churn         = churn_summary$n_churn,
  n_commits       = churn_summary$n_commits,
  mature_rate     = span_summary$mature_rate,
  mean_span_days  = span_summary$mean_span_days,
  n_authors       = span_summary$n_authors,
  seed            = seed
)
fwrite(out, "../../core/paper/outputs/lift_aiwork_helix.csv")
kable(out)
```
