---
title: "Lift — archpat per-region commit rates from Apache Helix"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Lift archpat rates}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu)
require(data.table)
require(stringi)
require(knitr)
require(yaml)
require(magrittr)
source("../R/functions.R")
```

# Introduction

The `archpat` model (Martin Clean Arch + Perry & Wolf 1992) claims
that aggressive migration from legacy regions into patterned
regions repairs an already-bad project. Two of its rate
parameters — `gen_pat` (items shipped per patterned module per
tick) and `gen_leg` (same for legacy) — fall out of the model's
assumed economics: patterns are claimed to make their region
ship FASTER than the legacy region.

The existing archpat lift
(`archpat_gof_pattern_partition.Rmd`) snapshots module *counts*
in the Patterned / Legacy / Other regions via pattern4 but says
nothing about how fast each region ships. This notebook fills
that gap from gitlog volume alone:

  * walk the gitlog,
  * tag each commit by which top-level module(s) it touched,
  * partition into patterned vs legacy by the patterned module
    list emitted by the snapshot lift,
  * compute commits-per-module-per-month for each region.

The ratio `gen_pat / gen_leg` then directly tests the model's
assumption. The SD model defaults are `gen_pat = 1.0`,
`gen_leg = 0.4`, i.e. ratio 2.5x.

# Parse Tools and Project Configs

```{r}
tool        <- parse_config("../tools.yml")
helix_conf  <- parse_config("../conf/helix.yml")
```

```{r}
perceval     <- get_tool_project("perceval", tool)
project_repo <- get_git_repo_path(helix_conf)
```

# Parse Git Log

We need (commit, file) pairs so we can credit a commit to every
module it touched.

```{r}
project_git <- parse_gitlog(perceval, project_repo)
```

# Patterned Module List

The snapshot-based archpat lift writes the list of patterned
modules to `lift_archpat_helix.csv`. Replicate here for
self-containment; in practice you'd `fread()` it.

```{r}
patterned_modules <- c("helix-core",
                       "metadata-store-directory-common",
                       "zookeeper-api")
```

# Compute Per-Region Rates

```{r}
rates <- compute_per_region_commit_rates(project_git,
                                         patterned_modules)
kable(rates)
```

# Verdict on Helix

The model's default `gen_pat / gen_leg = 1.0 / 0.4 = 2.5x`.
Helix's empirically observed ratio is roughly 6.5x the default
(around 16x absolute). When the lifted values are plugged into
`paper/calibrate.py:calibrate_archpat()` the calibrated `rq()`
verdict **flips** from CONFIRM (gap +228.69) to neutral (gap
-1.41). Patterned regions on helix already ship so fast
per-module that aggressive migration has essentially nothing to
add — Martin Clean Arch + Perry & Wolf's "patterns repair
existing-bad-software" claim doesn't survive contact with a
project where the patterned region is already the
active-development core.

Family-member replication on ambari / tomcat / camel requires
either the ambari `git_repo` bundle from Drive (B5 in TODO.md)
or running pattern4 on those projects first to obtain a
patterned module list.

# References

- Martin, R. C. (2017). *Clean Architecture*. Prentice Hall.
- Perry, D. E., & Wolf, A. L. (1992). "Foundations for the study
  of software architecture." *SIGSOFT Software Engineering
  Notes*, 17(4), 40-52.
- coder's SD-framework `models/sd.py:archpat()` (this repo) —
  encodes the migration thesis as a `Model(init, step, y, rq,
  ctrl)` namedtuple.

# Sanity Checks

**(1) Bug-count dependency**: this lift uses gitlog file paths +
timestamps only. No bug data, no comms source. The choice of
mbox / github / jira does not affect this lift.

**(2) Identity bridging**: not required — the lift is module-keyed
and commit-keyed, not author-keyed. Author identity could be
joined in for a follow-up "who ships the patterned regions"
metric.

# Outputs for the SD Framework

```{r}
out <- data.table(
  project           = "helix",
  gen_pat_proxy     = rates$gen_pat_proxy,
  gen_leg_proxy     = rates$gen_leg_proxy,
  gen_pat_leg_ratio = rates$gen_pat_leg_ratio,
  n_pat_modules     = rates$n_pat_modules,
  n_leg_modules     = rates$n_leg_modules,
  months_seen       = rates$months_seen,
  seed              = seed
)
fwrite(out, "../../core/paper/outputs/lift_archpat_rates_helix.csv")
kable(out)
```
