---
title: "Lift — Jira Issues and Comments (Helix, Camel)"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Lift Jira}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu); require(data.table); require(stringi)
require(knitr); require(yaml); require(magrittr); require(jsonlite)
source("../R/functions.R")
```

# Introduction

This notebook extracts Jira issues and their comments for the two
projects in the MYTHS family that ship with a Jira dump on disk:
Apache Helix and Apache Camel. Both dumps live under
`data/<project>/jira/issues/` and `data/<project>/jira/issue_comments/`.

The extraction uses `kaiaulu::parse_jira`, which returns a named list
of two data.tables — `issues` (one row per issue) and `comments`
(one row per comment, joinable to issues on `issue_key`).

Two known pitfalls flagged in `kaiaulu_notes/known_bugs.md`:

1. `parse_jira` does not return a flat table; it returns
   `list(issues, comments)`. Code that treats it as flat will silently
   break.
2. Filtering `issue_status` on `c("Closed","Resolved")` (as kaiaulu's
   own `metric.R` does) is broken on the Apache jira dump. The
   correct filter is `issue_status == "Done"`.
3. Installed `kaiaulu::parse_jira` takes a **directory** path
   (not a file) and chokes on `.DS_Store`.

The helper `lift_project_jira(project_conf)` in `R/functions.R`
encapsulates the per-project loop: stage every `*.json` under the
two configured directories into a clean tempdir (filtering out
sidecar files), run `parse_jira`, row-bind, and de-duplicate on
`issue_key` (issues) and `(issue_key, comment_id)` (comments).

# Configuration

```{r}
tool_conf  <- parse_config("../tools.yml")
helix_conf <- parse_config("../conf/helix.yml")
camel_conf <- parse_config("../conf/camel.yml")
```

# Extract Helix

```{r}
helix_jira <- lift_project_jira(helix_conf)
helix_jira$issues[,   project := "helix"]
helix_jira$comments[, project := "helix"]
cat("helix issues:  ", nrow(helix_jira$issues),   "\n")
cat("helix comments:", nrow(helix_jira$comments), "\n")
```

# Extract Camel

```{r}
camel_jira <- lift_project_jira(camel_conf)
camel_jira$issues[,   project := "camel"]
camel_jira$comments[, project := "camel"]
cat("camel issues:  ", nrow(camel_jira$issues),   "\n")
cat("camel comments:", nrow(camel_jira$comments), "\n")
```

# Summaries

## Issue-type distribution (Helix)

```{r}
kable(helix_jira$issues[, .N, by = issue_type][order(-N)])
```

## Issue-type distribution (Camel)

```{r}
kable(camel_jira$issues[, .N, by = issue_type][order(-N)])
```

## Issue-status distribution (Helix)

```{r}
kable(helix_jira$issues[, .N, by = issue_status][order(-N)])
```

## Issue-status distribution (Camel)

```{r}
kable(camel_jira$issues[, .N, by = issue_status][order(-N)])
```

## Bug share (informs lift_bugs, lift_brooksq, lift_rework)

```{r}
bug_share <- function(dt, label) {
  data.table(
    project   = label,
    n_issues  = nrow(dt),
    n_bugs    = nrow(dt[issue_type == "Bug"]),
    n_done    = nrow(dt[issue_status == "Done"]),
    n_done_bug = nrow(dt[issue_type == "Bug" & issue_status == "Done"])
  )
}
kable(rbind(
  bug_share(helix_jira$issues, "helix"),
  bug_share(camel_jira$issues, "camel")
))
```

# Write outputs

```{r}
dir.create("../../core/paper/outputs", showWarnings = FALSE, recursive = TRUE)

fwrite(helix_jira$issues,
       "../../core/paper/outputs/lift_jira_helix_issues.csv")
fwrite(helix_jira$comments,
       "../../core/paper/outputs/lift_jira_helix_comments.csv")
fwrite(camel_jira$issues,
       "../../core/paper/outputs/lift_jira_camel_issues.csv")
fwrite(camel_jira$comments,
       "../../core/paper/outputs/lift_jira_camel_comments.csv")
```

# Notes for downstream lifts

- Join key to gitlog: `commit_message_id_regex.issue_id`
  (`HELIX-[0-9]+`, `CAMEL-[0-9]+`). `parse_commit_message_id` writes
  the matched key into `commit_message_id`; that joins to
  `issue_key` here.
- `lift_brooksq` and `lift_rework` need `issue_type == "Bug"` rows.
- `lift_bugs` may also need `issue_status == "Done"` (NOT
  `c("Closed","Resolved")` — kaiaulu's `metric.R` bug).
- For communication-graph lifts (e.g. `lift_congruence` jira mode),
  `parse_jira_replies` consumes the same `parsed_jira` list shape
  this notebook produces.
